
---
title: "Search for future mining and landing sites on Mars"
author: Hadarou Sare

---

# Introduction

For more than decades,NASA and its partners has been developing technologies to extract precious metals and water from the Moon, Mars and other bodies in the solar system. The extraction of water is very important because it can reduce considerably the cost of launch since the water can be split into Hydrogen and Oxygen and use as propellant in space. That idea has been clearly supported recently by the US goverment and European partners with the Artemis and other projects. However, even though scientists and engineers have already developed the right technologies able to extract water and other resources from space, it is still unclear where they should land their technologies(Rover and others) and start mining the resources they are looking for. 

The goal of this project is to use the package "threejs" to visualize the future best mining sites on Mars(water/water ice/volatiles materials mining sites). The final product is a map showing locations of best landing sites and mining sites.
However, before getting into that step, I used some geological and engineering criteria to determine and selected the best landing/mining sites for the future rovers. The criteria used to select the landing/mining sites are: where are located minerals indicating the presence of water/water ice?, where do we have permanent sunlight to use as source of energy for processing the water and charging the rovers? Is the topography of the site suitable for the mobility of the rover to and from the processing center?... 
For this project, the most important criteria I  considered is the geological criteria for question of time. I choosed and explored 9 sites as training sites and then selected among those sites the best ones as future landing/mining sites.

# Materials and methods

I choosed 9 sites as training sites (Si polar swath"I00831006"; Lyot crater"V00823007"; Cerberus fossae"V00825005";Echus chasma"I00839002"; Sinus Meridiana west"I00849005"; Syrtis Major"I00853002"; Ophir planum"I00864002"; Marte vallis"I00868006; Eastern Meridiani"I00836002").
For each training site, I downloaded and used Raster data. Those data are Visible and infrared data in the “.TIFF” format and are combination of different band (9 0R 10 bands according to Mars Odysee website) and each band is showing data in a specific wavelength. I  then added all data (images) together to create a metadata containing all information from each site. For Mars, I don't need to resize any of the images because they all have the same size.I finally used the band ratio combination to determine the presence, abundance and distribution of water, water ice, ice or volatiles materials in each site I choose as training site and from there I was able to see which sites could be selected as the best future landing/mining sites. I finally get the GPS coordinates of those sites and plot them on a Martian globe using the package threejs.

# Data and Results
## PART1: Selection of the landing/mining sites

The goal here is to detect the site where we could have more water/ water ice/ volatiles materials. I applied some corrections to satelitte images I downloaded. For question of time I will only show what I did in only one site here but all sites I selected to be future landing/mining site are shown later in the second part of this project.

I loaded required packages

```{r, message=F, warning=F}
library(raster)
library(tidyverse)
library(sf)
library(rpart)
library(rpart.plot)
library(rasterVis)
library(mapedit)
library(mapview)
library(caret)
library(forcats)
library(sp)
library(raster)
```

I prepare the data by reading in the different bands that comprise the satelitte imagery. Each band refers to different spectrum.


```{r, message=F, warning=F}

band1 <- ("b01.tiff")
band2 <- ("b02.tiff")
band3 <- ("b03.tiff")
band4 <- ("b04.tiff")
band5 <- ("b05.tiff")
band6 <- ("b06.tiff")
band7 <- ("b07.tiff")
band8 <- ("b08.tiff")
band9 <- ("b09.tiff")
```



```{r, message=F, warning=F}
library(raster)

band1 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/b01.tiff")
band2 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/b02.tiff")
band3 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/b03.tiff")
band4 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/b04.tiff")
band5 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/b05.tiff")
band6 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/b06.tiff")
band7 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/b07.tiff")
band8 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/b08.tiff")
band9 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/b09.tiff")
```

Show result of band1

```{r}
print(band1)
```

Now, I check the size of image before combining them together using STACK to make sure the sizes match. The code used "res(bandX), where X is the number of the band" shows that all bands have the same size 20x20.


For image classification or image processing to map water or volatiles materials in selected sites, it is important to resize some image using for example the function "bandX <- aggregate(bandx, fact = 2)" in case they don't have the same size to make sure all images have the same size before adding them together using stack function in order to have a single data containg all informations from each band.
However, for Martian data and from the result after tests, we observed that all images have the same size so we will just add them all together later without resizing any image.


Now we can add all bands together to have a metadata containing all informations we are looking for and which will be helpful to continue the project.

Most of the time, when dealing with or Earth data, researchers usually  add only band2, band3 and band7 together because those wavelength are able to reflect the distribution of water.

However, we are dealing with the Martian data in our project and we observed that for Martian data, we have the same size so we will just add them all together.But I add only band4, band5, band6, band8 and band9 together because those wavelength are able to reflect the distribution of water on the selected site.
```{r}
image1 <- stack(Band2, Band3, Band7) #Here is how it works for Landsat data on Earth
#image1 <- stack(band4, band5, band6)
image2 <- stack(band4, band5, band6, band8, band9) #Here is for Mars images

```

Let's explore the images. Here, we are able to see the number of band used, the coordinate system the imagery is projected in, the resolution (or grid cell size) of the raster, and the resolution of the images


```{r}

#For Mars
nlayers(image2) #This give the number of bands use
crs(image2) #the coordinate system the imagery is projected in, and the resolution (or grid cell size) of the raster.
res #resolution of the images
```

##Results1
Now that we know a little more about the imagery we are using, let plot it. Since image is a multi-band raster,
we use the plotRGB function from the raster package, which allows us to specify what bands should be visualized.


###FALSE COLOR PLOT.The false color composite uses NIR (5) for red, red (4) for green, and green (3) for blue.


```{r}

par(col.axis="white",col.lab="white",tck=0)
plotRGB(image2, r = 5, g = 4, b = 3, axes = TRUE, stretch = "lin", main = "False Color Composite")
box(col="white")
```
###Indices of WATER

We can use Modified Normalized Difference Water Index [MNDWI=(Green-MIR)/ (Green+MIR)] to separate water from land. This idea is from Reference=https://www.researchgate.net/post/What_is_the_best_band_combination_for_highlighting_the_water_and_soil_together_in_LANDSAT_image2
The Author of the reference is Saygin Abdikan and is currently working at the Department of Geomatics Engineering, Hacettepe University. The Author researchGate is https://www.researchgate.net/profile/Saygin_Abdikan

The formula is as below:
[MNDWI=(Green-MIR)/ (Green+MIR)]
MNDWI <- (image[[2]] - image[[3]])/(image[[2]] + image[[3]]).

To plot the results with ggplot, we convert the raster into a data frame and use geom_tile.
The plot generated by the codes below show where are located water/ water ice/ volatiles materials the most and we have then know where to land the rovers and start the mining process.

Band4, Band5, and Band6 are use for Earth datas where Band4 represent the Green band in our case and Band6 represent the MIR band.

For Mars,the formula change little bit because all bands are collected and stored in the same file in Mars Odysee website. False color or black and white could reveal the distribution of water. However, we use the formula:

MNDWI <- (band4)/(1/band9).
This formula is derived after multiple simulations and observations

```{r}

#We derived This formula is derived after multiple observations and applied to different data before using it.
MNDWI <- (band4)/(1/band9)
#MNDWI <- (band4-band9)/(band4+band9)


as(MNDWI, "SpatialPixelsDataFrame") %>% 
  as.data.frame() %>%
  ggplot(data = .) +
  geom_tile(aes(x = x, y = y, fill = layer)) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "Water or Water ice or Volatiles materials", 
       x = " ", 
       y = " ") +
  scale_fill_gradient(high = "#CEE50E", 
                      low = "#087F28",
                      name = "Water or Water ice or Volatiles materials")
```

## PART2: Visualization of landing/mining sites on map
After detecting the sites that could be consider and selected as best future landing/mining site according to the figure above, We just get the geographical coordinates (locations from the onboard GPS of the spacecraft) of those sites and created an exel file with those coordinates and then plot them on Martian globe using the threejs package.

As seen above, the selection of the sites are firstly done by processing satellites images from Mars to see where the water or volatiles materials are abundant. In this section which is the second section of the project, we will get the geographical coordinates of those sites consider as best future landing sites and plot them on a Martian globe using the package threej. In fact, since the Martian gloge we will use in this project is already georeferenced with Martian coordiante system, then finding best future landing/mining site's geographical coordinates through our method in the first section will allow us to create an exel file where we will store those data (coordinate of the best future sites) and finally plot those sites on a Martian globe using the package threejs by calling the exel file in Rstudio.

I loaded required packages

```{r, message=F, warning=F}
library(threejs)
```

##Results2
I firstly prepared and loaded the data and then with the package called "streejs" I plot the exact locations of landing and mining sites on the Martian global map. I created a cvs file where I stored the latitudes and logitude of the future landing and mining sites selested and call that file in R.


```{r}
library(threejs)
library(csvread)
the_martian <- read.csv("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/visualize_of_landingMining_sites/the_martian.csv", comment.char="#")
                        
#the_martian.csv <- ("C:/Users/Owner/Desktop/sare  backup/c/user #hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/visualize_of_landi#ngMining_sites/the_martian.cvs")
  

globejs(img = "C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/visualize_of_landingMining_sites/Mars_Viking_MDIM21_ClrMosaic_global_1024.jpg", bodycolor="#555555", emissive="#444444",
        lightcolor="#555555",
        lat = the_martian$x,
        long = the_martian$y)

```



# Conclusions

From all sites I selected as training sites I choosed 9 sites as training sites (Si polar swath"I00831006"; Lyot crater"V00823007"; Cerberus fossae"V00825005";Echus chasma"I00839002"; Sinus Meridiana west"I00849005"; Syrtis Major"I00853002"; Ophir planum"I00864002"; Marte vallis"I00868006; Eastern Meridiani"I00836002"), Meridiana west seems to be the best one. The band ratio combination is used to determine the presence, abundance and distribution of water, water ice, ice or volatiles materials in each site before selecting some specific spots within Meridiana west as future landing/mining site.This is done using R language meaning that R is a suitable tool offering the user the choice to manipulate the data in the way he or she wants as well as to perfom diverses task on Earth and in space. R is definitely a great too that can help explorers, planetary scientists to overcome some difficulties they could not handle with other softwares.


# References

All sources are cited (to be updated later)
